---
title: "pine"
author: "rswaty"
date: "2021-03-17"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE}
pacman::p_load(tidyverse, ggsci, ggtext, readr, dyplr, DT)
theme_set(theme_bw())
```



## Visualizing the Pine data

First, the cleaned Pine data

```{r pineDT, echo=FALSE, message=FALSE, warning=FALSE}
pine <- read_csv("data/pine.csv")

datatable(pine)
```
<br>

Notes:

* Swaty added the "BURN_YEAR" column for ease of plotting
* Data is a compilation of spreadsheets supplied by Kristina Serbesoff-King (1996-2012) and Charles Oliver (2013-2020)

<br>

## Number of observations (rows) per year

```{r obsYear, echo=FALSE, message=FALSE, warning=FALSE}


yearObs <-
pine %>%
  count(MON_YEAR)

yearObsChart <-
  yearObs %>%
  ggplot(aes(x = MON_YEAR, y = n)) +
  geom_line(size = 1.5) +
  labs(x = "Year",
       y = "Number of rows",
       title = "Number of observations per year",
       caption = "Data from Florida Chapter")  +


  #Adjust the axes
  theme(plot.title = element_text(face="bold", size =14),
        axis.title.x = element_text(face="bold", size = 12),
        axis.title.y = element_text(face="bold", size = 12),
        legend.title = element_text(face="bold", size = 12))


yearObsChart
```


## Number of observations (rows) per year split out by Burn Unit (BU)

```{r message=FALSE, warning=FALSE, include=FALSE, results='hide'}
yearObsBU <- pine %>%
  count (MON_YEAR, BU)
```


```{r obsYearBU, echo=FALSE, message=FALSE, warning=FALSE}


# fix BU data type


yearObsBUChart <-
  yearObsBU %>%
  ggplot(aes(x = MON_YEAR, y = n, group =  BU, color = as.factor(BU))) +
  geom_line(size=1.5, position = position_jitter(width = 0.0001, height = 1.0)) +
  #geom_point(size =  2.0) +

  # My top color palette
  scale_color_simpsons(name = "BU") +
  # labels
  labs(x = "Year",
       y = "Number of rows",
       title = "Number of observations per year colored by BU",
       subtitle = "Lines jittered so that they would not totally overlap",
       caption = "Data from Florida Chapter")  +

  #Adjust the axes
  theme(plot.title = element_text(face="bold", size =14),
        axis.title.x = element_text(face="bold", size = 12),
        axis.title.y = element_text(face="bold", size = 12),
        legend.title = element_text(face="bold", size = 12))


yearObsBUChart
```


## Number of pine stems per plot

<br>

```{r message=FALSE, warning=FALSE, include=FALSE, results='hide'}
pine %>%
  mutate(MON_YEAR = as.character(MON_YEAR))
```


```{r numStems, echo=FALSE, message=FALSE, warning=FALSE}

stemsExplore <-
  pine %>%
  ggplot(aes(x = MON_YEAR, y = STEMS, group = MON_YEAR, fill = MON_YEAR)) +
  geom_boxplot(position = position_dodge(width = 0.9)) +
  geom_point(position = position_jitterdodge(seed = 1, dodge.width = 0.9)) +
  theme_bw() +
  labs(x = "Year",
       y = "Number of stemsp",
       title = "Total number of stems per plot",
       subtitle = "Dots jittered so that they would not totally overlap",
       caption = "Data from Florida Chapter")  +

  #Adjust the axes, remove legend
  theme(legend.position = "none",
        plot.title = element_text(face="bold", size =14),
        axis.title.x = element_text(face="bold", size = 12),
        axis.title.y = element_text(face="bold", size = 12))



stemsExplore


```


## Stem size per year

```{r message=FALSE, warning=FALSE, include=FALSE, results='hide'}

groupedPine <- pine %>%
  group_by(PRES,  SIZE,  MON_YEAR)%>%
  summarise(meanSTEMS = mean(STEMS))
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
stemsSizeYear <-groupedPine %>%

 ggplot(aes(x= MON_YEAR, y= meanSTEMS, group = SIZE, color = SIZE)) +
  geom_line(size=1.5) +
  geom_point(size =  2.0) +

  # My top color palette
  scale_color_simpsons() +

  # Theme
  theme_bw() +


  # Update the labels of the axes
  labs(x = "Year",
       y = "Mean number of stems per year at ABRP",
       title = "Summarizing number of stems across size categories",
       caption = "Data from Florida Chapter",
       color = "Stem Size")  +

  #Adjust the axes
  theme(plot.title = element_text(face="bold", size =14),
        axis.title.x = element_text(face="bold", size = 12),
        axis.title.y = element_text(face="bold", size = 12),
        legend.title = element_text(face="bold", size = 12))
stemsSizeYear
```


## Stem size per year split out by BU

```{r message=FALSE, warning=FALSE, include=FALSE, results='hide'}

groupedPineBU <- pine %>%
  group_by(PRES,  SIZE,  MON_YEAR, BU)%>%
  summarise(meanSTEMS = mean(STEMS))
```


```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=8}
groupedPineYearBU <-groupedPineBU %>%

  ggplot(aes(x= MON_YEAR, y= meanSTEMS, group = SIZE, color = SIZE)) +
  geom_line(size=1.0) +
  geom_point(size =  1.5) +
  facet_grid(rows = vars(BU), scales = "free") +
  # Update the labels of the axes
  labs(x = "Year",
       y = "Mean number of stems",
       title = "Stems per year by size class",
       subtitle = "Each facet is a BU; NOTE that scales are different") +


  # My top color palette
  scale_color_simpsons() +

  # Theme
  theme_bw()

  groupedPineYearBU
```

